Stellar Radiation is Critical for Regulating Star Formation and Driving Outflows in Low Mass Dwarf Galaxies

### 2018-07-18 ms.tex
Update on Overleaf.

\graphicspath{{./}{figures/}}

@@

\begin{abstract}
Ionizing radiation from recently formed massive stars is important in destroying cold, dense gas around new regions of star formation that would otherwise continue to form stars. Applying a local prescription for this pre-supernova feedback can be sufficient in matching star formation rates in a simulation with full radiative transfer on short timescales. However, the long-range effects of ionizing radiation are important in producing a volume-filling diffuse, warm-ionized medium in the ISM and in the disk-halo interface. This is essential for allowing supernovae to drive significant gas outflows that travel throughout and beyond the virial radius of the galaxy. A simulation with local radiation feedback only is unable to drive significant outflows as the supernovae are trapped well within the inner halo by higher density cold and warm neutral gas.
\end{abstract}

@@

\section{Introduction} \label{sec:intro}
Overview of importance of pre-SN feedback in regulating SF. Brief literature overview of recent simulations with full radiative transfer. Brief mention / discussion of current approximations / simplifications made in runs without RT that include something for radiation feedback. Motivate how it is unclear as to whether or not full fancy RT physics is necessary, or if 1) can just do SNe (unlikely), and 2) can just get by with local (e.g. Stromgren) approximation. How might this affect global galaxy properties beyond SFR?

Key here is that radiation affects role of SNe in driving galactic winds. This is known already (Peters2016, Hu2016, Hu2017, and more), but has not been demonstrated like this for outflowing properties, primarily b/c many methods use approximate effects of RT.

@@

\begin{figure*}
\centering
\includegraphics[width=0.49\linewidth]{sfr}
\includegraphics[width=0.49\linewidth]{mass}
\caption{}
\label{fig:sfr_mass_evolution}
\end{figure*}

@@

@@

Figure~\ref{fig:panel1}, at XX Myr, shows each just after the first few SN in each simulation. Already there are significant differences across runs. Gas outside the galaxy is warm ($\sim$~10$^{4}$~K) and ionized up to XX~pc above/below the plane of the disk in our fiducial run. This same gas is cold ($<10^4$~K) and neutral in both other runs. The contrast between the fiducial and shortrad radiation is seen most clearly in the ionized region in-plane and to the right of center in the top and bottom panels. This region contains massive stars that are capable of generating enough ionizing radiation to carve a channel out to the halo of our fiducial simulation; by construction, this does not occur in the shortrad case. The warm / ionized feature towards the center of each galaxy is gas from a recent SN. This is much more well contained and hotter in the noRT and shortrad runs than in our fiducial simulation.

\begin{figure*}
\centering
\includegraphics[width=0.99\linewidth]{DD0136_fiducial_shortrad_nort}
\caption{\textbf{In the final version of this figure I will add labels for each row marking the simulation, and a line giving the scale of each figure.}}
\label{fig:pane1l}
\end{figure*}

As these simulations evolve, the existence of these diffuse, ionized channels in the fiducial run easily allow significant outflows from SNe to escape the disk and move well into the halo of our galaxy. In contrast, the same SNe in the other two simulations are well contained, surrounded by shells of denser, neutral gas and unable to driving significant mass loss from the galaxy. In the noRT case, outflow does eventually develop, but it takes the factor of 5 increase in SFR for SN gas to finally break out from the neutral gas surrounding the galaxy.

\begin{figure*}
\centering
\includegraphics[width=0.99\linewidth]{DD0160_fiducial_shortrad_nort}
\caption{Panel}
\label{fig:pane2}
\end{figure*}

\section{Conclusion}  \label{sec:conclusion}
Very short. Just re-iterate points

\begin{thebibliography}{}

\end{thebibliography}

>>>>

\graphicspath{{./}{figures/}}
\usepackage{natbib}
\bibliographystyle{apj}

@@

\begin{abstract}
Placeholder
%{\bf Ignore: Ionizing radiation from recently formed massive stars is important in destroying cold, dense gas around new regions of star formation that would otherwise continue to form stars. Applying a local prescription for this pre-supernova feedback can be sufficient in matching star formation rates in a simulation with full radiative transfer on short timescales. However, the long-range effects of ionizing radiation are important in producing a volume-filling diffuse, warm-ionized medium in the ISM and in the disk-halo interface. This is essential for allowing supernovae to drive significant gas outflows that travel throughout and beyond the virial radius of the galaxy. A simulation with local radiation feedback only is unable to drive significant outflows as the supernovae are trapped well within the inner halo by higher density cold and warm neutral gas.}
\end{abstract}

@@

\section{Introduction} \label{sec:intro}
Historically, simulations of galaxy formation have suffered from the ``overcooling'' problem, whereby simulations including self-gravity and radiative cooling alone produce galaxies with far too many stars. This problem has been addressed by employing various models of feedback physics which produces self-regulating star formation in galaxies. Recently, it has become clear that simple models for stellar feedback, which account for supernovae (SNe) alone, are insufficient to fully regulate star formation and set galaxy properties consistent with observations (see \cite{NaabOstriker2017} for a recent review). Additional processes, such as stellar radiation, cosmic rays, or AGN, are often employed as additional sources of feedback. Modeling these processes in detail, however, is challenging.

Radiation from massive stars dominates the total feedback energy output of a stellar population \citep[e.g.][]{Lietherer1999,Agertz2013}, surpassing the energy ejection of supernova ($\sim 10^{51}$~erg) by two orders of magnitude. If the radiation couples effectively to the interstellar medium (ISM), it can be a substantial source of additional feedback. %Lower mass stars (size range) primarily drive photoelectric heating (cite) and H$_2$ dissociation through radiation in the far ultraviolet and Lyman-Werner bands, while more massive stars (range) emit significant amounts of ionizing radiation. 
Simulations including stellar radiation feedback followed through radiative transfer or radiation-hydrodynamics schemes have found it to be effective in regulating star formation and driving galactic winds \citep[e.g.][]{Wise2012,Kim2013,Sales2014,Rosdahl2015,Ocvirk2015,Pawlik2015,Peters2017}. This occurs for four reasons: 1) heating gas and preventing the formation of cold, dense star formation regions, 2) the destruction of cold, dense gas around recently formed stars, preventing further star formation, 3) momentum input by direct absorption of UV radiation by gas and (in some cases) dust through re-emission and scattering in the infrared, and 4) lowering the typical ISM densities in which SNe occur and greatly increasing their effectiveness. % the relevance / importance of RP from dust in the IR is beyond the scope of this work...

However, most works that employ stellar radiation feedback to account for these effects do so using various forms of sub-grid, approximate models to avoid the substantial additional cost of full radiative transfer. Many works use a Stromgren approximation whereby the particles / cells within the Stromgren radius a radiating star are heated and kept ionized, with additional approximations made to account for overlapping ionized regions \citep[e.g.][]{HQM2011,Hu2016,Hu2017}. Other works employ some form of energy or momentum injection localized to the region immediately around a star particle \citep[e.g.][(\textit{need more})]{Agertz2013,Roskar2014,Ceverino2014,Forbes2015}. Although some of these approximate methods account for long-range effects of diffuse radiation \citep{HQM2012,Hopkins2018} most cases treat local radiation {\it only}, confined to energy or momentum injection in a limited physical region around a star particle. It is unclear what, if any, effect the long-range ionization and heating processes play in regulating star formation and driving galactic winds. Although the approximate radiation feedback models are capable of driving significant galactic outflows from galaxies, it is unclear if they capture the full effects of stellar ionizing radiation.

In this work we present a comparison of three simulations of the evolution of an isolated, low mass dwarf galaxy that account for detailed stellar feedback. Our fiducial model uses an adaptive ray-tracing radiative transfer method to follow stellar ionizing radiation. We test its effectiveness in regulating star formation and driving galactic winds by comparison to a simulation without any ionizing radiation and a simulation with only localized radiation around each star. We discuss our methods in Section~\ref{sec:methods} and present our results in Section~\ref{sec:results}.

@@

\begin{figure*}
\centering
\includegraphics[width=0.49\linewidth]{sfr}
\includegraphics[width=0.49\linewidth]{mass}
\caption{The star formation rate (left) and gas and stellar mass (right) of each of our three simulations over time.}
\label{fig:sfr_mass_evolution}
\end{figure*}

@@

\begin{figure}
\centering
\includegraphics[width=0.98\linewidth]{metal_retention}
\caption{The fraction of metals contained within the ISM of each galaxy (blue) and the fraction ejected beyond the virial radius (orange). As in Figure~\ref{fig:sfr_mass_evolution}, the fiducial run is given as solid lines, while the noRT and shortrad are dash-dotted and dashed respectively.}
\label{fig:metal_retention}
\end{figure}

@@

Figure~\ref{fig:panel1}, at 17~Myr, shows each just after the first few SN in each simulation. Already there are significant differences across runs. Gas outside the galaxy is warm ($\sim$~10$^{4}$~K) and ionized up to $\sim$500~pc above/below the plane of the disk in our fiducial run. This same gas is cold ($<10^4$~K) and neutral in both other runs. The contrast between the fiducial and shortrad radiation is seen most clearly in the ionized region in-plane and to the right of center in the top and bottom panels. This region contains massive stars that are capable of generating enough ionizing radiation to carve a channel out to the halo of our fiducial simulation; by construction, this does not occur in the shortrad case. The warm / ionized feature towards the center of each galaxy is gas from a recent SN. This is much more well contained and hotter in the noRT and shortrad runs than in our fiducial simulation.

\begin{figure*}
\centering
\includegraphics[width=0.99\linewidth]{DD0136_fiducial_shortrad_nort}
\caption{\textbf{In the final version of this figure I will add labels for each row marking the simulation, and a line giving the scale of each figure.} Edge-on slices of each simulation showing gas number density (left), temperature (middle), and hydrogen ionization fraction (right) 17~Myr after the formation of the first star in each run.}
\label{fig:panel1}
\end{figure*}

As these simulations evolve, the existence of these diffuse, ionized channels in the fiducial run easily allow significant outflows from SNe to escape the disk and move well into the halo of our galaxy. In contrast, the same SNe in the other two simulations are well contained, surrounded by shells of denser, neutral gas and unable to driving significant mass loss from the galaxy. In the noRT case, outflow does eventually develop, but it takes the factor of 5 increase in SFR for SN gas to finally break out from the neutral gas surrounding the galaxy. This is shown more clearly in Figure~\ref{fig:panel2}, which gives each simulation 40~Myr after the formation of the first stars. Comparing the shortrad and fiducial runs long term (\textit{give sentence on how these panels look 200 + Myr down the line but don't actually show this}).

\begin{figure*}
\centering
\includegraphics[width=0.99\linewidth]{DD0160_fiducial_shortrad_nort}
\caption{Same as Figure~\ref{fig:panel1}, but at 40~Myr.}
\label{fig:panel2}
\end{figure*}

\section{Conclusion}  \label{sec:conclusion}
In agreement with previous works, we find that stellar radiation is an important source of feedback that helps regulate star formation in dwarf galaxies. (more)

\bibliography{msbib} % if your bibtex file is called example.bib
